---
title: "2022 PIPS Exam"
subtitle: 2022 Programming in Psychological Science Exam
output:
  pdf_document:
    toc: no
  word_document:
    toc: no
  html_document:
    toc: no
---

<!-- Authors: Michael D. Nunez, Hannes Rosenbusch-->

## Exam instructions

Place your ID clearly visible on your right side. Read these exam instructions carefully.

This exam must be completed individually. You **are** allowed to use your notes, R, Python, and websites **that do not have interactive chat features**. You can only use your laptops, pens, and paper. Cellphone usage is not allowed. **Watching videos (e.g. YouTube) is not allowed.** Your computer speakers must be muted and no headphones are allowed. **You cannot communicate with anyone in any form during the exam** (e.g. no Slack, posting on Stack Overflow, morse code, etc.). If you do so, you will automatically fail this exam.

**Pick any 12** of the following 20 problems to submit on Canvas. You can attempt more than 12 problems, but **you may only submit 12 answers**. Note for the first 16 questions you must use **R** and for the last 4 questions you must use **Python**. You may only use the R libraries and Python modules specified. Some questions have more specific rules about what libraries/modules to use. R questions and Python questions are labeled with a "R" and "P" respectively.

You must include the name of each problem (e.g "E.R2") in the Canvas submission before your answer. Your answer can be copied and pasted from a R or Python script. But you should **not** submit an entire .R or .py script. **You must also return these 2 front-back printed exam papers with your name and student ID.** Please enter the provided exam number on Canvas. Submit the Canvas Exam and return these papers by 16:00. 

On request of the examiner (or his/her representative) you should be able to identify yourself with a valid ID card. If you have any technical problems or questions about the exam, raise your hand.

## Write your name:

## Write your student ID:

## Place this exam number on Canvas:
```{r, eval=TRUE, echo=FALSE}
sample(1:10^9, 1)
```

## You may only use these libraries for the R questions starting on the next page:
```{r, eval=FALSE}
library(titanic)
library(ggplot2)
library(dplyr)
```
\newpage

## E.R1
What are the names of the first two arguments used by the function median()? 

## E.R2
Define a function named `is_tensor()` that takes one argument, this function returns true if the input array has more than 2 dimensions and false if the input array has 2 dimensions or less. The function should also produce an error that says "`Input must be an array`" if the input is not an array. You may only use base R functions.

## E.R3
Using the `titanic_train` data set from the library `titanic`, create a function `was_there_a` that takes one argument `search_term`. This function should return a logical value indicating whether there was a male passenger on the Titanic with the `search_term` being part of their name. You may only use base R functions.

## E.R4
Create a array of size 132 by 5 by 360 with random draws from the standard normal distribution in every element using `rnorm()` with default arguments. Now index the 101st to 343rd element of the last dimension of your array to return 243 matrices within a 3 dimensional array. Do everything without using any loops and only base R functions.

## E.R5
Load the `titanic_train` data set from the `titanic` package into your workspace. 
What is the percentage of the passengers with a *known* age was under 30 years old?

## E.R6
In terms of **R style** what are two things wrong with the following code? Rewrite the code to fix the problems you identified with the **R style**. You may not use lintr, nor Styler in R Studio.
```{r, eval=FALSE}
# This line samples 1000 random normal variables and then adds that to 1000 beta
# variables with shape1 = 1 and shape2 = 2 parameters. Then it takes the 
# absolute value of that sum. Then it takes the square root of the sum. Then it 
# takes the mean of that sum. Then assigns it to the variable "mean_data"
mean_data <- mean(sqrt(abs(rnorm(1000) + rbeta(1000, shape1 = 1, shape2 = 2))))
```

## E.R7
Using the built-in `ChickWeight` data set, display the weights of the chickens at time point 21 in a violin plot. Note that you may (but do not have to) use a function from the `ggplot2` library. How many chickens do not have a value for this time point?

## E.R8
Create the following `data.frame` without typing out the full vector of any of the columns.
```{r, eval=TRUE, echo=FALSE}
data.frame(year = seq(-60, 0, by=20),
           people = rep(c('Roman'),4),
           government = rep(c('republic', 'empire'), each=2))
```

## E.R9
Load the `titanic_train` data set from the `titanic` package into your workspace. Add a new logical variable called `leo` to the data frame.  This variable has the value `TRUE` when the person was in third class and did not survive. For all other cases it has the value `FALSE`. Print how many passengers have a `TRUE` value the `leo` column. You may use base R and/or `dplyr` functions.


## E.R10
Only using base R functions, make the following vector without writing all the texts/numbers: 
`c("Student1", "Student2", "Student3", "Student4", "Student5")`
Then use a base R function to change the substring `"Student"` to `"Master"` for each element in the vector.

## E.R11
```{r, eval=FALSE}
hist(sqrt(abs(rnorm(1000))))
```
Convert this R code into something that uses at least 4 pipes (either `dplyr` pipes or base R pipes).

## E.R12
Your friend is writing code to calculate the amount of times throwing 3 dice will contain only 5s. Your friend is using 1000 simulated dice rolls to do so. However their code always returns 0 rolls as 5-5-5.

What is wrong with their code below? Can you fix their code to better simulate the number of 5-5-5 throws? You should be able to share the code with your friend and get the exact same answer.

```{r, eval=FALSE}
num_rolls <- 1000
all_five <- 0
for (index in 1:num_rolls) {
  set.seed(1234)
  dice_roll <- sample(1:6, 3, replace = TRUE)
  if (all(dice_roll == 5)) {
    all_five <- all_five + 1
  }
}
```

## E.R13
Load the `titanic_train` data set from the `titanic` package into your workspace. Using `dplyr` functions (also known as `dplyr` "verbs"), return a data frame with only 3rd class passengers that lists passenger fares in descending order.

## E.R14
```{r, eval=FALSE}
set.seed(11)
x = rnorm(20)
r = rnorm(20)
y = 1 + x + r
```
Make a scatter plot **only using base R** that plots variable `x` on the x-axis and `y` on the y-axis. Label your x-axis "Independent variable" and your y-axis "Dependent variable". **`ggplot2` functions are not allowed for this question**.

## E.R15
Create a while loop in which you draw 100 new random normal distribution draws (using the standard normal distributions, e.g. default parameters) until the minimum is less than -5. Print `Positive outlier found!` if the maximum is greater than 5 in the loop. Use only base R functions. You must use an "explicit loop".


## E.R16
In terms of **R style** what are two things wrong with the following code? Rewrite the code to fix the problems you identified with the **R style**. You may not use lintr, nor Styler in R Studio.
```{r, eval=FALSE}
set.seed(11)
x = rnorm(20)
r = rnorm(20)
y = 1 + x + r
```

<!-- Python -->
## You may only use these modules for the following Python questions:
```{python, eval=FALSE}
import numpy as np
import pandas as pd
import matplotlib.pyplot as plt
```

## E.P1

Define a **Python** **function** that takes two vectors as inputs. The **function** then plots a scatter plot of the first input on the x-axis and the second input on the y-axis. Use the `matplotlib` library to make your plots.

## E.P2

Using **Python**, make the following `pandas.DataFrame` without typing out the full vector of any of the columns.

```{python, eval=TRUE, echo=FALSE}
import numpy as np
import pandas as pd
all_data = {"year" : np.arange(-60, 1, step=20),
            "people" : np.repeat(np.array(['Roman']),4),
           "government" : 
             np.repeat(np.array(['republic', 'empire']), 2)} 
my_data = pd.DataFrame(all_data)
print(my_data)
```

## E.P3
Write code that reads the csv file at this location using `pandas` in **Python**:

https://tinyurl.com/pipsTitanic

Write a for loop over the rows (passengers) of the data file's original order.
Whenever the current passenger is female and survived, **and also** the previous passenger was male, print: "`I'll never let go, Jack. I'll never let go. I promise.`"

## E.P4

Write a function in **Python** with one argument `start_num`. This function counts up by 0.1 when given a number `start_num` greater than 0, and counts down by one whole number when given a number `start_num` less than 0. The function should stop counting when it reaches 100 or -100. The function should also print "`I'm not good at counting backwards`" before it counts backwards, but not when it counts forwards.






